A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching

### 2016-11-01 JTB_ms/JTB_ms.Rmd
sigmaCS=0.9, kROS=80

@@

```{r load, include=FALSE}
# Source metrics reported in text
source("R/reported_metrics.R")
```

@@

| Symbol  | Description                            | Units               | Value   |
| ------- | -------------------------------------- | ------------------- | ------- |
| $n_{NH}$ | N:C molar ratio in host biomass | -- | 0.19 |
| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | 0.2 |
| $n_{NX}$ | N:C molar ratio in prey biomass | -- | 0.13 |
| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | 0.03 |
| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | 0.03 |
| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | -- | 0.9 |
| $\sigma_{CH}$ | Proportion host carbon turnover recycled | -- | 0.9 |
| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | -- | 0.9 |
| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | -- | 0.9 |
| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | 0.1292 |
| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | 20e-6 |
| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | 0.048 |
| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | 0.46e-6 |
| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | 10 |
| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | 1 |
| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | 0.1 |
| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | 1.34 |
| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | 112 |
| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | 40-80 |
| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | 2.8 |
| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | 0.25 |
| $b$ | Scaling parameter for strength of bleaching response | -- | 5 |

@@

An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is made available to the photosynthesis SU (scaled to the relative amount of symbiont biomass), according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}

>>>>

  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10% of metabolic CO2 produced is passively available to photosynthesis, to reflect the fact that that ~10% of total internal DIC at pH 7 is in the form of CO2. probably actually less than this since internal pH may be closer to 7.5. This does not change any of the figures substantively except the hysteresis figure--overshoot no longer occurs with specific parameters run. may need to just change light range or rate of change to reproduce overshoot.

@@

```{r load, include=FALSE}
# Source metrics reported in text
source("R/reported_metrics.R")
# Source default parameter values
source("R/def_pars.R")
defpars <- def_pars()
```

@@

| Symbol  | Description                            | Units               | Value   |
| ------- | -------------------------------------- | ------------------- | ------- |
| $n_{NH}$ | N:C molar ratio in host biomass | -- | $`r defpars$nNH`$ |
| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | $`r defpars$nNS`$ |
| $n_{NX}$ | N:C molar ratio in prey biomass | -- | $`r defpars$nNX`$ |
| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHT0`$ |
| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | $`r defpars$jST0`$ |
| $\sigma_{NH}$ | Proportion host N-turnover recycled to host | -- | $`r defpars$sigmaNH`$ |
| $\sigma_{CH}$ | Proportion host C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCH`$ |
| $\sigma_{NS}$ | Proportion symbiont N-turnover recycled to symbiont | -- | $`r defpars$sigmaNS`$ |
| $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | $`r defpars$jXm`$ |
| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | $`r defpars$jNm`$ |
| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | $`r defpars$KN`$ |
| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | $`r defpars$kCO2`$ |
| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHGm`$ |
| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | $`r defpars$yCL`$ |
| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | $`r defpars$astar`$ |
| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | $`r defpars$kNPQ`$ |
| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | $`r defpars$kROS`$ |
| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | $`r defpars$jCPm`$ |
| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | $`r defpars$jSGm`$ |
| $b$ | Scaling parameter for strength of bleaching response | -- | $`r defpars$b`$ |

@@

An additional fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively made available to the photosynthesis SU, according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}

### 2016-11-01 JTB_ms/JTB_ms.tex
sigmaCS=0.9, kROS=80

  \begin{abstract}
  Coral reef ecosystems owe their ecological success--and vulnerability to
  climate change--to the symbiotic metabolism of corals and
  \emph{Symbiodinium} spp. The urgency to understand and predict the
  stability and breakdown of these symbioses (i.e., coral `bleaching')
  calls for the development and application of theoretical tools. While
  the biological complexity of the symbiosis is well-suited to dynamical
  modeling, the theoretical complexity of certain approaches makes them
  inaccessible to many biologists, thus limiting their potential
  application and utility. Here, we develop a simplified dynamic
  bioenergetic model of coral-algal symbiosis that demonstrates realistic
  steady-state behavior across gradients of light, nutrients, and feeding.
  Furthermore, through a mechanistic treatment of photo-oxidative stress,
  the model displays dynamic responses of bleaching and recovery, and
  suggests that these phenomena represent transitions between alternative
  stable states. A suite of complex observed behaviors reproduced by this
  model suggests it has captured many important attributes of the system;
  meanwhile, its modular framework and open source R code are designed to
  facilitate expansion and problem-specific applications. We see high
  utility for this model in elucidating the mechanistic underpinnings of
  complex responses, generating testable hypotheses, and predicting
  integrated responses to simultaneous changes in multiple environmental
  inputs.
  \end{abstract}

@@

0.19

@@

0.2

@@

0.13

@@

0.03

@@

Proportion host nitrogen turnover recycled

@@

0.9

@@

Proportion host carbon turnover recycled

@@

0.9

@@

Proportion symbiont nitrogen turnover recycled

@@

Proportion symbiont carbon turnover recycled

@@

0.1292

@@

20e-6

@@

0.048

@@

0.46e-6

@@

@@

1

@@

0.1

@@

1.34

@@

112

@@

40-80

@@

2.8

@@

0.25

@@

@@

An additional fixed proportion of metabolic CO\textsubscript{2}
generated by host biomass turnover is made available to the
photosynthesis SU (scaled to the relative amount of symbiont biomass),
according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \cdot {S \over H} \end{equation}

>>>>

  \begin{abstract}
  Coral reef ecosystems owe their ecological success--and vulnerability to
  climate change--to the symbiotic metabolism of corals and
  \emph{Symbiodinium} spp. The urgency to understand and predict the
  stability and breakdown of these symbioses (i.e., coral `bleaching')
  calls for the development and application of theoretical tools. While
  the biological complexity of the symbiosis is well-suited to dynamical
  modeling, the theoretical complexity of certain approaches makes them
  inaccessible to many biologists, thus limiting their potential
  application and utility. Here, we develop a simplified dynamic
  bioenergetic model of coral-algal symbiosis that demonstrates realistic
  steady-state patterns in coral growth and symbiont abundance across
  gradients of light, nutrients, and feeding. Furthermore, through a
  mechanistic treatment of photo-oxidative stress, the model displays
  dynamic responses of bleaching and recovery, and suggests that these
  phenomena represent transitions between alternative stable states. A
  suite of complex observed behaviors reproduced by the model suggests it
  captures many important attributes of the system; meanwhile, its modular
  framework and open source R code are designed to facilitate expansion
  and problem-specific applications. We see high utility for this modeling
  framework in generating testable hypotheses and predicting integrated,
  mechanistic responses of corals to environmental change.

  Latest development--sigmaCH and sigmaCS changed to 0.1 meaning 10\% of
  metabolic CO2 produced is passively available to photosynthesis, to
  reflect the fact that that \textasciitilde{}10\% of total internal DIC
  at pH 7 is in the form of CO2. probably actually less than this since
  internal pH may be closer to 7.5. This does not change any of the
  figures substantively except the hysteresis figure--overshoot no longer
  occurs with specific parameters run. may need to just change light range
  or rate of change to reproduce overshoot.
  \end{abstract}

@@

\(0.18\)

@@

\(0.13\)

@@

\(0.2\)

@@

\(0.03\)

@@

Proportion host N-turnover recycled to host

@@

\(0.9\)

@@

Proportion host C-turnover recycled to photosynthesis

@@

\(0.1\)

@@

Proportion symbiont N-turnover recycled to symbiont

@@

Proportion symbiont C-turnover recycled to photosynthesis

@@

\(0.13\)

@@

\(1.95\times 10^{-5}\)

@@

\(0.035\)

@@

\(1.5\times 10^{-6}\)

@@

\(10\)

@@

\(1\)

@@

\(0.1\)

@@

\(1.34\)

@@

\(112\)

@@

\(80\)

@@

\(2.8\)

@@

\(0.25\)

@@

\(5\)

@@

An additional fixed proportion of metabolic CO\textsubscript{2}
generated by host biomass turnover is passively made available to the
photosynthesis SU, according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}

### 2016-11-01 JTB_supp/JTB_supp.Rmd
sigmaCS=0.9, kROS=80

* $\sigma_{NH}, \sigma_{CH}, \sigma_{NS}, \sigma_{CS}$: Proportions of nitrogen and carbon recycled from host and symbiont biomass turnover
    + Assuming highly efficient recycling to avoid loss to the environment:
    + Values used: `r with(defpars, c(sigmaNH, sigmaCH, sigmaNS, sigmaCS))` (dimensionless)

@@

# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environments
env1 <- init_env(time=time, L=c(20,40,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
env2 <- init_env(time=time, L=c(20,40,2), N=c(1e-6,1e-6,0), X=c(1e-5,1e-5,0))

# Run simulations
ss1 <- with(run_coral_ss(env=list(L=30,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
ss2 <- with(run_coral_ss(env=list(L=30,N=1e-6,X=1e-5), pars=defpars), last(S$S/H$H))
run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, "kCO2", 6))#
plot_run(run)
```

>>>>

* $\sigma_{NH}, \sigma_{NS}$: Proportions of nitrogen recycled from host and symbiont biomass turnover
    + Assume highly efficient recycling by both symbiont and host to avoid loss to the environment:
    + Values used: `r with(defpars, c(sigmaNH, sigmaNS))` (dimensionless)
* $\sigma_{CS}$: Proportion of metabolic CO~2~ produced by symbiont recycled passively to photosynthesis
    + Assume that CO~2~ produced within symbiont cells is recycled with high efficiency, potentially aided by symbiont carbon concentrating mechanisms
    + Value used: 0.9 (dimensionless)
* $\sigma_{CH}$: Proportions of metabolic CO~2~ produced by host recycled passively to photosynthesis 
    + Assume that most metabolic CO~2~ produced within host cells rapidly converts to HCO~3~^-^ at intracellular pH ~ 7.2 [@Venn:2009hy], and that only ~10% actually remains as CO~2~ [@Zeebe:2001wz] that could diffusively supply photosynthesis. This passive supply is differentiated from all active host carbon concentrating mechanisms (CCMs) that facilitate CO~2~ supply to photosynthesis, which are separately treated collectively as the $jCO~2~$ flux.
    + Values used: 0.1 (dimensionless)

@@

# Set run time vector
time <- seq(1, 365, 0.1)  # if single model run, use time input

# Initialize environments
env1 <- init_env(time=time, L=c(22,44,2), N=c(1e-7,1e-7,0), X=c(1e-5,1e-5,0))
env2 <- init_env(time=time, L=c(22,44,2), N=c(1e-5,1e-5,0), X=c(1e-5,1e-5,0))

# Run simulations
ss1 <- with(run_coral_ss(env=list(L=33,N=1e-7,X=1e-5), pars=defpars), last(S$S/H$H))
run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", ss1))
ss2 <- with(run_coral_ss(env=list(L=33,N=1e-5,X=1e-5), pars=defpars), last(S$S/H$H))
run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", ss2))

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=defpars)
plot_run(run)
```
